\begin{acknowledgement}
Time really flies, and two glorious years seem in the blink of an eye! It is always a good time in Hong Kong.~``Life is like a box of chocolates. You never know what you're gonna get''~(Forrest Gump, 1994). After my undergraduate in Jilin, China, I never imagined having the good fortune of studying in Hong Kong. At this moment, full of gratitude, I sincerely appreciate every one of my supervisors and labmates around me.

First and foremost, I would like to express the depth of my gratitude to all the professors who ever supervised, supported, helped, taught, and encouraged me. It seems impossible for me to get into academia without their patient guidance, constant support and encouragement, insightful and helpful discussions, and invaluable inspiration. In specific, I would like to thank Professor Shiqi Wang for his supervision of my research and study. Professor Shiqi Wang is a rigorous and meticulous researcher with a charming, humble, discreet, and diligent profile. His rigorous research philosophy inspires me. He takes a serious attitude to his students' research and paperwork. I am tasked to perform very rigorous experiments and write high-quality paperwork with strict guidelines and the highest standard. Since then, research rigor has been deeply rooted in my research work and personal life. Also, I would like to show my appreciation to Professor Richard Yi-Da Xu for teaching me machine learning and optimization theory and offering me an opportunity to continue my study, and Professor William Hsu for enlightening me and enhancing my confidence in my future career path. Last but not least, I would like to express my sincerest thanks to Professor Antoni B.~Chan and Professor Ka-Chun Wong for their excellent teaching.

Next, I would like to deliver my appreciation to my dear friends and amazing labmates around me: Zhijian Hou, Xiaoyuan Zhang, Raphael Elikplim Nkrow, Xin Chen, Zhaojun Ding, Dingquan Li, Tao Ni, Ziquan Liu, Bangyao Wu, Hao Zhang, Bin Zhu, Chen Chen, Tianlu Ma, Lenong Chu, Yifan Zhang, Wei Wu, Wei Lin, Siyu Ren, Qiangqiang Wu, Peilin Chen, Xuelin Shen, Shurun Wang, Keyan Ding, Yixuan Li, Yingwen Zhang, Chris Xing Tian, Binzhe Li, Jiancong Chen, Qiudan Zhang, Haoyuan Wang, Yuhao Liu, Tao Chen, Thanh Nguyen Huu, Andrea Musumeci, Andres Ruiz Serrano, Jeffry Oktavianus, Muhammad Masood, and Boey Kai Zhe, and thank their help during my stay in Hong Kong.

I am particularly grateful to my roommates, Dr.~Xin Chen and Dr.~Zhaojun Ding, for their generous help and assistance, especially during the tough times of COVID-19 Pandemic (2020-2023). I also would like to give my grateful appreciation to Dr.~Zhijian Hou for his personality influence,~\eg, self-discipline, concentration on research, motivation, rational thinking, and diligence, and Dr.~Xiaoyuan Zhang for his enlightenment and broader vision. Apart from them, a special appreciation should be shown to Dr.~Dingquan Li for his guidance and inspiration on my paperwork, Dr.~Darcy Bird, who has helped me a lot with my English writing since high school, and Miss Louisa Tian for her accompany in Hong Kong.

Furthermore, I wish to express my gratitude to my parents, Mrs.~Wenying Qin and Mr.~Wenjun Jia, and my sister, Dr.~Fei Jia, for their endless love, support, motivation, and encouragement.

Last but not least, to the unexpected and unforgettable in Hong Kong, and find the courage to beyond boundaries.
\newline

\noindent Shuyue Jia (Bruce S.~Jia)\\
\noindent Workplace~\#11, Room M5006, 5/F, Run Run Shaw Creative Media Center (CMC), City University of Hong Kong, 18 Tat Hong Avenue, Kowloon, Hong Kong\\
\noindent December $31^{\mathrm{st}}$ (New Year's Eve), 2022

\end{acknowledgement}